Calendar-based graphics for visualizing people's daily schedules

### 2018-10-10 archive/rjournal/main.Rmd
archived rjournal version

---
title: Calendar-based graphics for visualizing people's daily schedules
author:
  - name: Earo Wang
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics,
    - Monash University, VIC 3800
    - Australia
    email: earo.wang@monash.edu
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics,
    - Monash University, VIC 3800
    - Australia
    email:  dicook@monash.edu
  - name: Rob J Hyndman
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics,
    - Monash University, VIC 3800
    - Australia
    email:  rob.hyndman@monash.edu
abstract: >
  Calendars are broadly used in society to display temporal information, and events. This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout. The function \code{frame\_calendar} uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts. The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city. Faceting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the R package \pkg{sugrrants}.
output: 
  rticles::rjournal_article:
  includes:
    in_header: preamble.tex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \pkg{ggplot2} [@R-ggplot2]. The data restructuring approach is consistent with the tidy data principles available in the \pkg{tidyverse} [@R-tidyverse] suite. The methods are implemented in a new package called \pkg{sugrrants} [@R-sugrrants] .

@@

The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.  It is also related to recent work by @R-geofacet which provides methods in the \pkg{geofacet} package to arrange data plots into a grid, while preserving the geographical position. Both of these show data in a spatial context.

>>>>

---
title: Calendar-based graphics for visualizing people's daily schedules
author:
  - name: Earo Wang
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics,
    - Monash University, VIC 3800
    - Australia
    email: earo.wang@monash.edu
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics,
    - Monash University, VIC 3800
    - Australia
    email:  dicook@monash.edu
  - name: Rob J Hyndman
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics,
    - Monash University, VIC 3800
    - Australia
    email:  rob.hyndman@monash.edu
abstract: >
  Calendars are broadly used in society to display temporal information, and events. This paper describes a new R package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout. The function \code{frame\_calendar} uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts. The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city. Faceting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the R package \pkg{sugrrants}.
output: 
  rticles::rjournal_article:
  includes:
    in_header: preamble.tex
bibliography: ["RJreferences.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \CRANpkg{ggplot2} [@R-ggplot2]. The data restructuring approach is consistent with the tidy data principles available in the \CRANpkg{tidyverse} [@R-tidyverse] suite. The methods are implemented in a new package called \CRANpkg{sugrrants} [@R-sugrrants] .

@@

The work is inspired by @Wickham2012glyph, which uses linear algebra to display spatio-temporal data as glyphs on maps.  It is also related to recent work by @R-geofacet which provides methods in the \CRANpkg{geofacet} package to arrange data plots into a grid, while preserving the geographical position. Both of these show data in a spatial context.

### 2018-10-10 archive/rjournal/main.tex
archived rjournal version

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, giving a restructuring of the
data, that can then be piped into grammar of graphics definitions of
plots, as used in \pkg{ggplot2} (Wickham et al. 2018). The data
restructuring approach is consistent with the tidy data principles
available in the \pkg{tidyverse} (Wickham 2017) suite. The methods are
implemented in a new package called \pkg{sugrrants} (Wang, Cook, and
Hyndman 2018) .

@@

The work is inspired by Wickham et al. (2012), which uses linear algebra
to display spatio-temporal data as glyphs on maps. It is also related to
recent work by Hafen (2018) which provides methods in the \pkg{geofacet}
package to arrange data plots into a grid, while preserving the
geographical position. Both of these show data in a spatial context.

@@

\bibliography{references.bib,packages.bib}

>>>>

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, giving a restructuring of the
data, that can then be piped into grammar of graphics definitions of
plots, as used in \CRANpkg{ggplot2} (Wickham et al. 2018). The data
restructuring approach is consistent with the tidy data principles
available in the \CRANpkg{tidyverse} (Wickham 2017) suite. The methods
are implemented in a new package called \CRANpkg{sugrrants} (Wang, Cook,
and Hyndman 2018) .

@@

The work is inspired by Wickham et al. (2012), which uses linear algebra
to display spatio-temporal data as glyphs on maps. It is also related to
recent work by Hafen (2018) which provides methods in the
\CRANpkg{geofacet} package to arrange data plots into a grid, while
preserving the geographical position. Both of these show data in a
spatial context.

@@

\bibliography{RJreferences.bib}
